---
title: "DataDownload"
author: "Garret Wais"
date: "02/08/2015"
output: html_document
---

# See Project-1.R for required packages and to run program.
# This file, DataDownload_andProcessing.Rmd, shows all R processing steps.

### R session Used:
```{r session info}
sessionInfo()
```


### Download file
```{r download file}
fileurl = "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
```

### Store file in temp directory
```{r store file in temp}
tempdir = tempdir()
```

### Create temporary zip file
```{r creat temp dir}
temp <- tempfile(tmpdir = tempdir, fileext = ".zip")
```

### Download file into temp directory and zip
```{r download data}
download.file(fileurl, destfile = temp, method = "wget")
```

### List file
```{r file list}
file <- unzip(temp, list=TRUE)
file
```

### Create 'data' folder in current working director
```{r create data folder}
if (!file.exists("data")){dir.create("data")}
```

### Create a folder to hold Final PNG files
```{r}
if (!file.exists("final_plot_pngs")){dir.create("final_plot_pngs")}
```

### Unzip and store data
```{r unzip data}
unzip(temp, exdir = "./data", overwrite = TRUE)
```

### Unlink temprorary file and directory
```{r unlink temps}
unlink(temp)
unlink(tempdir)
```

### Assign variable name 'file' to data file
```{r assign file variable}
datafile <- c("./data/household_power_consumption.txt")

#decided not to use the following, but maybe helpful for future use.
#data_subset <- read.csv.sql(file, header = TRUE, sep=";", sql = "select * from file where (Date == '1/2/2007' OR Date == '2/2/2007')" )
```

### Load data as data.frame
```{r load data}
housedata <- read.table(datafile, header = TRUE, sep = ";", na.strings = "?")
```

### Omit NAs
```{r omit NAs}
housedata <-na.omit(housedata)
```

### Subset by specified time period
```{r subset data to time period}
housedata <- housedata[housedata$Date %in% c("1/2/2007", "2/2/2007"),]
```

### Convert 'Date' field to date format
```{r convert to date}
housedata$Date <- strptime(paste(housedata$Date,housedata$Time), "%d/%m/%Y %H:%M:%S")
```
### Data has been processed. Please see; plot1.Rmd, plot2.Rmd, plot3.Rmd, and plot4.Rmd for their respective code. Thank you.